The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-07-18 WavesData-2019/LatexNotes/notes.tex
NM

% Basic editing
\newcommand{\tocite}{{\color{blue}(to cite)}}
\newcommand{\vsp}[1]{\vspace{#1 pc} \noindent}
\newcommand{\np}{\newpage \noindent}
% Derivatives
\newcommand{\pd}[2]    { \frac{\partial #1} {\partial #2} }
\newcommand{\ppd}[2]  { \frac{\partial^2 #1}{{\partial #2}^2} }
\newcommand{\pdi}[2] { {\partial_#2} #1 }
\newcommand{\td}[2] { \frac{d #1} { d #2 } }
\newcommand{\grad}{\nabla}
\newcommand {\Lap} {\grad^2}
% Vectors and operators
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\norm}[1]{\left\| #1 \right\|}
\newcommand{\mean}[1]{\left< #1 \right>}
\newcommand{\eps}{\varepsilon}
\newcommand{\defeq}{\mathrel{\mathop:}=}

@@

% Dashed integral
\def\Xint#1{\mathchoice
   {\XXint\displaystyle\textstyle{#1}}%
   {\XXint\textstyle\scriptstyle{#1}}%
   {\XXint\scriptstyle\scriptscriptstyle{#1}}%
   {\XXint\scriptscriptstyle\scriptscriptstyle{#1}}%
   \!\int}
\def\XXint#1#2#3{{\setbox0=\hbox{$#1{#2#3}{\int}$}
     \vcenter{\hbox{$#2#3$}}\kern-.5\wd0}}
\def\ddashint{\Xint=}
\def\dashint{\Xint-}
\newcommand{\intt}{\dashint}%_0^{2 \pi}}
\newcommand{\dx}{\, dx}

@@

\author{
C. Tyler Bolles\thanks{University of Michigan},
Andrew J.~Majda\thanks{Courant Institute of Mathematical Sciences}, 
M.~N.~J.~Moore\thanks{Florida State University}, 
Di Qi\thanks{Courant Institute of Mathematical Sciences} }
\maketitle

@@

We measure the state variable $u$ as the displacement of the surface from equilibrium, so that $\mean{u} = 0$. Consequently, the momentum of $u$ vanishes
\begin{equation}
\Mo[u] \defeq \int_{-\pi}^{\pi} u \dx = 0
\end{equation}
Next, the definition of the characteristic amplitude in \eqref{ampdefn} implies that $\mean{u^2} = \pi^{-1}$. Consequently, the energy of $u$ is fixed as
\begin{equation}
\En[u] \defeq \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = 1
\end{equation}

@@

The statistical matching condition is
\begin{equation}
\meanup{\Hdn} = \meandn{\Hdn}
\end{equation}
An experimental observation is that the incoming skewness is small, $\meanup{H_3} \approx 0$. Inserting and simplifying gives
\begin{equation}
\label{H3H2}
\frac{\meandn{H_3^+}} {\meandn{H_2^+} - \meanup{H_2^+}} = \frac{C_2}{C_3} \dratdn^2
\end{equation}
%where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
To convert to dimensional, i.e.~experimental, values we first note that
\begin{align}
\mean{H_3^+} = \frac{\pi}{3} \mean{u^3} = 
\frac{1}{3} \pi^{-1/2} \frac{\mean{\eta^3}}{\etastd^3} = \frac{1}{3} \pi^{-1/2} \skw(\eta)
\end{align}
Second, to convert the downstream surface slope, we note
\begin{align}
\pd{u}{\tilde{x}} = \frac{\lamfac \lamdn}{\pi^{3/2} \etastd} \pd{\eta}{\xi}
\end{align}
where slope with respect to $\xi$ is the same as with respect to dimensional $x$.
Note: I believe it is important here to use $\lamdn$ since we are matching $\Hdn$.
Hence, we have
\begin{align}
&\mean{H_2^+} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
\left( \frac{\lamfac \lamdn}{\pi \etastd} \right)^2 \var(\eta_x)
\end{align}
%
Then algebra gives
\begin{equation}
\frac{\meandn{H_3}} {\meandn{H_2} - \meanup{H_2}} = 
\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2 \lamdn^2} 
\left( \frac{\skwdn(\eta)} { \vardn(\eta_x) - \varup(\eta_x)} \right)
\end{equation}
Also we have the simplification
\begin{equation}
\frac{C_2}{C_3} = \frac{\pi^{3/2} \delup^2}{9 \lamfac^2 \epsup}
\end{equation}
Hence, \eqref{H3H2} gives
\begin{equation}
\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^4
\end{equation}
In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
% Note: previously I had D^2 on the RHS. The power of 4 comes from using lamdn consistently in the nondimensionalization.

>>>>

% Basic editing
\newcommand{\tocite}{{\color{blue}(to cite)}}
\newcommand{\vsp}[1]{\vspace{#1 pc} \noindent}
\newcommand{\np}{\newpage \noindent}
% Derivatives
\newcommand{\pd}[2]    { \frac{\partial #1} {\partial #2} }
\newcommand{\ppd}[2]  { \frac{\partial^2 #1}{{\partial #2}^2} }
\newcommand{\pdi}[2] { {\partial_#2} #1 }
\newcommand{\td}[2] { \frac{d #1} { d #2 } }
\newcommand{\grad}{\nabla}
\newcommand {\Lap} {\grad^2}
% Vectors and operators
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\norm}[1]{\left\| #1 \right\|}
\newcommand{\mean}[1]{\left< #1 \right>}
\newcommand{\eps}{\varepsilon}
\newcommand{\dx}{\, dx}

%\newcommand{\defeq}{\mathrel{\mathop:}=}

@@

@@

\author{
C.~Tyler Bolles\thanks{University of Michigan},
Andrew J.~Majda\thanks{Courant Institute of Mathematical Sciences}, 
M.~N.~J.~Moore\thanks{Florida State University}, 
Di Qi\thanks{Courant Institute of Mathematical Sciences} }
\maketitle

@@

We measure the state variable $u$ as the displacement of the surface from equilibrium, so that $\mean{u} = 0$. Consequently, the momentum of $u$ vanishes
\begin{equation}
\Mo[u] \equiv \int_{-\pi}^{\pi} u \dx = 0
\end{equation}
Next, the definition of the characteristic amplitude in \eqref{ampdefn} implies that $\mean{u^2} = \pi^{-1}$. Consequently, the energy of $u$ is fixed as
\begin{equation}
\En[u] \equiv \frac{1}{2} \int_{-\pi}^{\pi} u^2 \dx = 1
\end{equation}

@@

The statistical matching condition is
\begin{equation}
\meanup{\Hdn} = \meandn{\Hdn}
\end{equation}
An experimental observation is that the incoming skewness is small, $\meanup{H_3} \approx 0$. Inserting and simplifying gives
\begin{equation}
\label{H3H2}
\frac{\meandn{H_3^+}} {\meandn{H_2^+} - \meanup{H_2^+}} = \frac{C_2}{C_3} \dratdn^2
\end{equation}
%where $\Delta \mean{H_2} =  \meandn{H_2} - \meanup{H_2}$  indicates the difference between the upstream and downstream values.
To convert to dimensional, i.e.~experimental, values we first note that
\begin{align}
\mean{H_3^+} = \frac{\pi}{3} \mean{u^3} = 
\frac{1}{3} \pi^{-1/2} \frac{\mean{\eta^3}}{\etastd^3} = \frac{1}{3} \pi^{-1/2} \skw(\eta)
\end{align}
Second, to convert the downstream surface slope, we note
\begin{align}
\pd{u}{\tilde{x}} = \frac{\lamfac \lamdn}{\pi^{3/2} \etastd} \pd{\eta}{\xi}
\end{align}
where slope with respect to $\xi$ is the same as with respect to dimensional $x$.
Note: I believe it is important here to use $\lamdn$ since we are matching $\Hdn$.
Hence, we have
\begin{align}
&\mean{H_2^+} = \pi \mean{ \left( \pd{u}{\tilde{x}} \right)^2} = 
\left( \frac{\lamfac \lamdn}{\pi \etastd} \right)^2 \var(\eta_x)
\end{align}
%
Then algebra gives
\begin{equation}
\frac{\meandn{H_3}} {\meandn{H_2} - \meanup{H_2}} = 
\frac{\pi^{3/2} \, \etastd^2} {3 \lamfac^2 \lamdn^2} 
\left( \frac{\skwdn(\eta)} { \vardn(\eta_x) - \varup(\eta_x)} \right)
\end{equation}
Also we have the simplification
\begin{equation}
\frac{C_2}{C_3} = \frac{\pi^{3/2} \delup^2}{9 \lamfac^2 \epsup}
\end{equation}
Hence, \eqref{H3H2} gives
\begin{equation}
\frac{\skwdn(\eta)} {\vardn(\eta_x) - \varup(\eta_x)} = \frac{1}{3} \epsup^{-3} \dratdn^3
\end{equation}
In particular, the ratio on the left, which can be measured in the experiments, is predicted to scale as the inverse cube of the wave amplitude and as the square of the depth ratio.
% Note: previously I had D^2 on the RHS. The power of 3 comes from using lamdn consistently in the nondimensionalization.
